The reference model approach in feature selection problems \vspace{.1in}

### 2019-05-27 paper/ref_approach.tex
improved introduction

\documentclass[american,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath,amsthm}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Reference model in feature selection problems},
            pdfauthor={Federico Pavone, Juho Piironen, Aki Vehtari},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true,
            colorlinks=true,
            linkcolor=black,
            citecolor=RoyalBlue,
            filecolor=black,
            urlcolor=RoyalBlue,
          }
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=american]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[variant=american]{english}
\fi
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{placeins}
\usepackage{graphicx,grffile,subcaption}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{3}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\title{Reference model in feature selection problems 
	\vspace{.1in}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Federico Pavone, Juho Piironen, Paul-Christian B\"{u}rkner and Aki Vehtari}

@@

In statistical applications, one of the main steps of the modelling workflow is covariate or feature selection, that can be seen as a special case of model reduction. From a full Bayesian perspective, when the goal of the analysis is to do prediction, no selection should be carried out and integration over all parameters uncertainties, i.e. posterior distributions, is the recommended procedure. However, a high dimensional parameter space brings different difficulties as computation burden and overfitting. For such reason different feature selection procedures have been proposed both in the frequentist and Bayesian setting (e.g. see \cite{paper:feature_selection}, \cite{paper:vehtari_ojanen} and \cite{paper:model_selection}). Sometimes the task of the analysis is not prediction, but rather inference on which variables are significantly related to the phenomenon of interest. A common case is microarray data in medicine or biology, where for example for a very large number of genes, expression levels are measured for two sets of patients labeled as \textit{positive} or \textit{negative} to some disease (typically a \textit{small n large p} scenario).  The main goal of the analysis is to spot which subset of these genes is significantly related to the disease in order to perform further experiments. A possible way to proceed is to perform hypothesis testing to assess the significance of each gene, based on some modelling assumptions. \cite{book:efron} refers to this as multiple hypothesis testing problem. Such example could be still considered as feature selection, intended this time as the problem of ``identifying \textit{all} features that are predictive about (that is, statistically related to) the target variable'' \citep{paper:projpred}.

Regardless which kind of selection problem we are trying to solve, we consider examples of the reference model approach all those methods that make use of a full-encompassing model, i.e. the reference one. Such an idea is not new to the model reduction literature, it has been used for long time using different names. \cite{book:harrell} refers to the reference model as a full model that can be thought as the ``gold standard'' and shows some examples of how it can be used to look for a sparser approximation. He highlights some of the benefits as the possibility to calculate the accuracy with which the submodel approximates the best model and the inheritance of the shrinkage when it is applied to the full model. \cite{paper:faraggi_nn} deal with the necessity of identifying interpretable risk groups in the context of survival data. Neural networks typically perform very well in terms of prediction, but it is not very clear how to interpret the significance of each covariate. \cite{paper:faraggi_nn} propose to combine the benefits of the neural networks, which become what we call reference models, with the interpretability of regression trees, in order to provide a framework to make inferences about risk groups. \cite{paper:paul_preconditioning}, using the term preconditioning, explore approximating models fitting Lasso \citep{paper:tibshirani_lasso} or stepwise regression against consistent estimates $\hat{y}$ of a reference model instead of the observed response quantities $y$. \cite{book:harrell} includes also the possibility of using simple least squares for the approximating models as long as the target is the linear predictor of a full regression model \hl{CHECK}.
\\
The reference model approach has been used for long time also in the Bayesian framework, an example the pioneering work by \cite{paper:reference_lindley}. In the case of generalised linear models (GLMs), \cite{paper:tran_predictivelasso} provide a procedure to estimate and do variable selection with the goal of prediction. The best approximating model is recovered by optimising the Kullback-Liebler divergence between the predictive distribution of the reference model and the approximating model plus an $l_{1}$ penalty on the submodel coefficients. The resulting problem is convex and thus many efficient optimisation algorithm are available. Following \cite{paper:goutis_projection}, \cite{paper:original_proj}  instead project the posterior distribution of the reference model in each candidate model and then select the smallest one with enough explanatory power avoiding any prior elicitation for the submodels. In the special case of candidate models in the family of the GLMs, the optimisation problem correspond to the maximum likelihood estimation problem having the expected model point prediction by the reference model instead of the observed target values. Different extensions have followed \cite{paper:original_proj}, as \cite{paper:nott_projection} that add an $l_{1}$ penalty to select the best approximating model. \cite{paper:projpred} extend the projection algorithm with the clustered projection and choose the best model comparing the predictive performance with the reference model one.

\cite{paper:projpred} show in their experiments the benefits of the projection predictive approach in terms of predictive ability of the selected model and smaller size compared to other methods. It could be of interest to know how often a certain method includes non-relevant features. One could argue that such question can not have a general answer, because it depends both on the data and, when it is used, on the choice of the reference model. Moreover in real settings it is hard to define what relevant means, in most of the cases all the features have some level of relevance in term of prediction ability and thus the question looses of meaning.  However, it is still possible to simulate certain artificial scenarios and control which features are included. \cite{paper:model_selection} report results about the proportion of relevant and non-relevant features selected with different variable selection methods, showing overall better results in favour of the projection predictive approach. The explored submodels using the projection show also a higher performance and stability with respect to the compared methods, given the same submodel size.

In this paper we try to give further examples and intuitions on why the reference model helps. Using a real world dataset, we show the improvements in the selection when using the projection predictive approach versus the stepwise regression. However it has been used, the reference model can be always seen acting as a filter on the observed data. Following this idea, we devise a simple reference model approach that can be applied on top of any selection procedure. Our goal is to show the reference model benefits, regardless of what specific procedure is applied. We take as reference problem the normal means estimation problem. We apply different state-of-the-art methods comparing in each case the results using the data or using a reference model. In case of full Bayesian methods, we report results in terms of shrinkage and accuracy of the posterior estimates, whereas when a full selection is carried out, we compute quantities as false discovery rate, sensitivity and a measure of the stability of the selection. We believe that these results should motivate the use of a reference model when doing model reduction, regardless of the specific method used and, thus, we hope this work can be of interest to a large number of practitioners.

@@

\text{SESE}&=\sum_{j=1}^{p}\mathbb{E}_{\theta_{j}}[(\theta_{j}-\theta^{0}_{j})^{2}|\mathbf{z}] \label{eq:SESE}

>>>>

\documentclass[american,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath,amsthm}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={The reference model approach in feature selection problems},
            pdfauthor={Federico Pavone, Juho Piironen, Paul-Christian B\"{u}rkner, Aki Vehtari},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true,
            colorlinks=true,
            linkcolor=black,
            citecolor=RoyalBlue,
            filecolor=black,
            urlcolor=RoyalBlue,
          }
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=american]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[variant=american]{english}
\fi
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{placeins}
\usepackage{graphicx,grffile,subcaption}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{3}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\title{The reference model approach in feature selection problems 
	\vspace{.1in}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Federico Pavone, Juho Piironen, Paul-Christian B\"{u}rkner and Aki Vehtari}

@@

In statistical applications, one of the main steps of the modelling workflow is covariate or feature selection, which can be seen as a special case of model reduction. From a full Bayesian perspective, when the goal of the analysis is to do prediction, no selection should be carried out and integration over all parameters uncertainties, i.e. posterior distributions, is the recommended procedure. In a high dimensional parameter space the choice of the prior plays an even more key role due to different difficulties that can rise, as computation burden and overfitting. In addition to that, sometimes collecting all the features for a future observation is not possible or too expensive. In order to solve or, at least, simplify these problems, different feature selection procedures have been proposed both in the frequentist and the Bayesian setting \cite[e.g. see][]{paper:feature_selection,paper:vehtari_ojanen,paper:model_selection}. Sometimes the task of the analysis is not prediction, but rather inference on which variables are substantially related to the phenomenon of interest. This typically happens in a \textit{small n large p} scenario, as in many microarray data analysis in medicine or biology. A very large number of genes is usually analysed in two groups of individuals labeled as \textit{positive} or \textit{negative} to some disease. The main goal of the analysis is to spot the subset of explanatory genes in order to perform further experiments \cite[see examples in][]{book:efron}.

Either the target of the selection is the minimal subset of feature with good enough predictive ability, or the whole subset of relevant variables, the selection process is going to be based on assumptions regarding the ``true'' data generation mechanism. Making assumptions means making models, thus we identify two families of feature selection, or more in general model reduction, approaches depending on how the data generation mechanism is modelled: the data based and the reference model based ones. The former includes all those methods that use the observed empirical distribution as model approximation of the data generation mechanism, as for example happens in the Lasso selection \cite{paper:tibshirani_lasso} or in the stepwise backward/forward regression, where the selected submodels balance sparsity with predictive ability on the observed data. The reference model based methods use instead the predictive distribution of a full-encompassing model, i.e. the reference model, as an approximation of the data generation mechanism. Such an idea is widely present in the literature with different names. For example, \cite{book:harrell} refers to the reference model as a full model that can be thought as the ``gold standard'' and shows some examples of how it can be used to look for a sparser approximation. He highlights some of the benefits as the possibility to calculate the accuracy with which the submodel approximates the best model and the inheritance of the shrinkage when it is applied to the full model. \cite{paper:faraggi_nn} deal with the necessity of identifying interpretable risk groups in the context of survival data. Neural networks typically perform very well in terms of prediction, but it is not very clear how to interpret the significance of each covariate. \cite{paper:faraggi_nn} propose to combine the benefits of the neural networks, which become what we call reference models, with the interpretability of regression trees, in order to provide a framework to make inferences about risk groups. \cite{paper:paul_preconditioning}, using the term preconditioning, explore approximating models fitting Lasso \citep{paper:tibshirani_lasso} or stepwise regression against consistent estimates $\hat{y}$ of a reference model instead of the observed response quantities $y$. \cite{book:harrell} includes also the possibility of using simple least squares for the approximating models as long as the target is the linear predictor of a full regression model \hl{CHECK}.
The reference model approach has been used for long time also in the Bayesian framework, an example the pioneering work by \cite{paper:reference_lindley}. In the case of generalised linear models (GLMs), \cite{paper:tran_predictivelasso} provide a procedure to estimate and do variable selection with the goal of prediction. The best approximating model is recovered by optimising the Kullback-Liebler divergence between the predictive distribution of the reference model and the approximating model plus an $l_{1}$ penalty on the submodel coefficients. The resulting problem is convex and thus many efficient optimisation algorithm are available. Following \cite{paper:goutis_projection}, \cite{paper:original_proj}  instead project the posterior distribution of the reference model in each candidate model and then select the smallest one with enough explanatory power avoiding any prior elicitation for the submodels. In the special case of candidate models in the family of the GLMs, the optimisation problem correspond to the maximum likelihood estimation problem having the expected model point prediction by the reference model instead of the observed target values. Different extensions have followed \cite{paper:original_proj}, as \cite{paper:nott_projection} that add an $l_{1}$ penalty to select the best approximating model. \cite{paper:projpred} extend the projection algorithm with the clustered projection and choose the best model comparing the predictive performance with the reference model one.

%\cite{paper:projpred} show in their experiments the benefits of the projection predictive approach in terms of predictive ability of the selected model and smaller size compared to other methods. It could be of interest to know how often a certain method includes non-relevant features. One could argue that such question can not have a general answer, because it depends both on the data and, when it is used, on the choice of the reference model. Moreover in real settings it is hard to define what relevant means, in most of the cases all the features have some level of relevance in term of prediction ability and thus the question looses of meaning.  However, it is still possible to simulate certain artificial scenarios and control which features are included. \cite{paper:model_selection} report results about the proportion of relevant and non-relevant features selected with different variable selection methods, showing overall better results in favour of the projection predictive approach. The explored submodels using the projection show also a higher performance and stability with respect to the compared methods, given the same submodel size.

In this paper we try to bring motivations for the use of a reference model. A properly designed model is able to clean part of the noise present in the data, thus to provide an improved and more stable selection process. As an example, we show the improvements in the selection when using the projection predictive approach versus the stepwise regression with a real world dataset. We argue that however the reference model has been used, it can be always seen acting as a filter on the observed data. Following this idea, we devise a simple reference model approach that can be applied on top of any selection procedure. We thus show the reference model benefits, regardless of what specific procedure is applied, using as a benchmark the normal means estimation problem. We apply different state-of-the-art methods comparing in each case the results basing the selection on the data or on the reference model. In case of full Bayesian methods, we report results in terms of shrinkage and accuracy of the posterior estimates, whereas when a full selection is carried out, we compute quantities as false discovery rate, sensitivity and a measure of the stability of the selection. We believe that these results show how the core reason why the reference model based methods perform well is the reference model itself, rather than the specific way of using it. Of course we recognise that all these methods diversify how the subset of variables is actually chosen and which properties it has, thus it remains fundamental to investigate and develop such different procedures. Therefore we hope that this work can increase the interest on this family of methods and it can be of interest to a large number of practitioners.

@@

\text{SESE}&=\sum_{j=1}^{p}\mathbb{E}_{\theta_{j}}[(\theta_{j}-\theta^{0}_{j})^{2}|z_{1},..,z_{p}] \label{eq:SESE}
